\Huge
\noindent {\bf \textsc{Preface}}\\
\large
\emph{A Note to Students, Teachers, and other Readers}
\vspace{1in}
\normalsize

This text is used in a mathematical software course at VMI that provides an introduction to Matlab and Mathcad.  However, it is also intended to be a \emph{course} book instead of an all inclusive resource.  I encourage my students to take full advantage of the built-in help capabilities of these software packages, additional texts (I keep a few in a small library that is always available to students) and to use Google - I certainly did when learning the material and still do today. \\ 

This text is much shorter than a traditional text.  I have tried to provide the right amount of discussion and examples.  The exercise sets have a small number of problems, but I assign all of them when I teach the course.  I have tried to give problems that come from real-world situations and contemporary data and aren't simply contrived just to fill space.  Some of the problems reference the electronic course management system at VMI (Angel), but these can be ignored as the data sets are also included in the Appendix. \\

This text is also an ``open'' text.  If you wish to change the text for your needs, please do so.  I would also share source files if you are interested.  The Creative Commons copyright must be honored and that any changes be acknowledged and that the resulting work be used only in non-commercial areas.\\

Note that throughout the text, I make reference to colors (regarding text or pictures).  This book is printed in black and white on purpose, but you can contact me for the full color pdf version (www.vmi.edu/math $>$ faculty).\\

I welcome any comments or corrections.\\

Sincerely,\\

Troy Siemers\\
\indent troy@vmi.edu



